Selection games and Arhangelskii's convergence principles

### 2019-10-30 articles/selection-alpha-games/article.tex
several selection/arhangelskii updates

\begin{definition}
Say a collection \(\mc A\) is \term{sequence-like} if it satisfies the following
for each \(A\in\mc A\).
\begin{itemize}
\item \(|A|\geq\aleph_0\).
\item If \(A'\subseteq A\) and \(|A'|\geq\aleph_0\), then \(A'\in\mc A\).
\end{itemize}
\end{definition}

@@

It follows that \(\Gamma_X,\Gamma_{X,x}\) are both sequence-like. We also require
the following.

\begin{definition}
Say a collection \(\mc A\) is \term{almost-sequence-like} if
for each \(A\in\mc A\), there is \(A'\subseteq A\) such that:
\begin{itemize}
\item \(|A'|=\aleph_0\).
\item If \(A''\) is a cofinite subset of \(A'\), then \(A''\in\mc A\).
\end{itemize}
\end{definition}

So all sequence-like sets are almost-sequence-like.

\begin{theorem}
Let \(\mc B\) be sequence-like. Then \(\alpha_1(\mc A,\mc B)\) holds if and only
if \(\plI\notprewin G_{cf}(\mc A,\mc B)\).
\end{theorem}

@@

\begin{theorem}
Let \(\mc A\) be almost-sequence-like and \(\mc B\) be sequence-like. 
Then \(\alpha_2(\mc A,\mc B)\) holds if and only
if \(\plI\notprewin G_1(\mc A,\mc B)\).
\end{theorem}

@@

\begin{theorem}
Let \(\mc A\) be almost-sequence-like and \(\mc B\) be sequence-like. 
Then \(\alpha_4(\mc A,\mc B)\) holds if and only
if \(\plI\notprewin G_{<2}(\mc A,\mc B)\) if and only if
\(\plI\notprewin G_{fin}(\mc A,\mc B)\).
\end{theorem}

@@

\begin{theorem}
Let \(\mc B\) be sequence-like. Then \(\plI\prewin G_{<2}(\mc A,\mc B)\)
if and only if \(\plI\prewin G_{fin}(\mc A,\mc B)\).
\end{theorem}

\begin{proof}
Assume \(\bigcup\mc A\) is well-ordered.
Given a winning predetermined strategy \(A_n\) for \(\plI\) in
\(G_{<2}(\mc A,\mc B)\), consider \(F_n\in[A_n]^{<\aleph_0}\). We set
\[
  F_n^*
=
  \begin{cases}
    \emptyset
      & \text{ if }
    F_n\setminus\bigcup\{F_m:m<n\}=\emptyset
      \\
    \{\min(
      F_n\setminus\bigcup\{F_m:m<n\}
    )\}
      & \text{ otherwise}
  \end{cases}
\]
Since \(|F_n|<2\), we have that \(\bigcup\{F_n^*:n<\omega\}\not\in\mc B\).
In the case that \(\bigcup\{F_n^*:n<\omega\}\) is finite, we immediately
see that \(\bigcup\{F_n:n<\omega\}\) is also finite and therefore not
in \(\mc B\). Otherwise \(\bigcup\{F_n^*:n<\omega\}\not\in\mc B\)
is an infinite subset of \(\bigcup\{F_n:n<\omega\}\), and thus
\(\bigcup\{F_n:n<\omega\}\not\in\mc B\) too. Therefore
\(A_n\) is a winning predetermined strategy for \(\plI\) in
\(G_{fin}(\mc A,\mc B)\) as well.
\end{proof}

\begin{theorem}
Let \(\mc B\) be sequence-like. Then \(\plI\win G_{<2}(\mc A,\mc B)\)
if and only if \(\plI\win G_{fin}(\mc A,\mc B)\).
\end{theorem}

@@

\begin{theorem}
Let \(\mc B\) be sequence-like. Then \(\plII\win G_{<2}(\mc A,\mc B)\)
if and only if \(\plII\win G_{fin}(\mc A,\mc B)\).
\end{theorem}

\begin{theorem}
Let \(\mc B\) be sequence-like. Then \(\plII\markwin G_{<2}(\mc A,\mc B)\)
if and only if \(\plII\markwin G_{fin}(\mc A,\mc B)\).
\end{theorem}

\begin{theorem}\label{pawlikowskii}
Let \(\mc A\) be almost-sequence-like and \(\mc B\) be sequence-like.
Then \(\plI\win G_{fin}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{fin}(\mc A,\mc B)\).
\end{theorem}

\begin{proof}
We assume \(\plI\win G_{fin}(\mc A,\mc B)\).
As \(\mc A\) is almost-sequence-like, there is a strategy \(\sigma\)
witnessing \(\plI\win G_{fin}(\mc A,\mc B)\) where
\(|\sigma(s)|=\aleph_0\) and \(\sigma(s)\cap\bigcup\ran{s}=\emptyset\)
(that is, \(\sigma\) never replays the choices of \(\plII\))
for all partial plays \(s\) by \(\plII\).

For each \(s\in\omega^{<\omega}\), suppose \(F_{s\rest m}\in[\bigcup A]^{<\aleph_0}\)
is defined for each \(0<m\leq|s|\).
Then let \(s^\star:|s|\to[\bigcup\mc A]^{<\aleph_0}\) be defined by
\(s^\star(m)=F_{s\rest m+1}\), and define \(\tau':\omega^{<\omega}\to\mc A\)
by \(\tau'(s)=\sigma(s^\star)\). Finally, set 
\([\sigma(s^\star)]^{<\aleph_0}=\{F_{s\concat\tuple{n}}:n<\omega\}\), and
for some bijection \(b:\omega^{<\omega}\to\omega\) let \(\tau(n)=\tau'(b(n))\)
be a predetermined strategy for \(\plI\) in \(G_{fin}(\mc A,\mc B)\).

Suppose \(\alpha\) is a counterattack by \(\plII\) against \(\tau\), so 
\[
  \alpha(n)
    \in
  [\tau(n)]^{<\aleph_0}
    =
  [\tau'(b(n))]^{<\aleph_0}
    =
  [\sigma(b(n)^\star)]^{<\aleph_0}
\]
It follows that \(\alpha(n)=F_{b(n)\concat\tuple{m}}\) for some \(m<\omega\).
In particular, there is some infinite subset \(W\subseteq\omega\) and \(f\in\omega^\omega\)
such that \(\{\alpha(n):n\in W\}=\{F_{f\rest n+1}:n<\omega\}\).
Note here that \((f\rest n+1)^\star=(f\rest n)^\star\concat\tuple{F_{f\rest n+1}}\).
This shows that \(F_{f\rest n+1}\in[\sigma((f\rest n)^\star)]^{<\aleph_0}\) 
is an attempt by \(\plII\) to defeat \(\sigma\), which fails. Thus 
\(\bigcup\{F_{f\rest n+1}:n<\omega\}=\bigcup\{\alpha(n):n\in W\}\not\in\mc B\),
and since this set is infinite (as \(\sigma\) prevents \(\plII\)
from repeating choices) we have \(\bigcup\{\alpha(n):n<\omega\}\not\in\mc B\) too.
Therefore \(\tau\) is winning.
\end{proof}

\begin{proposition}\label{auto-asl}
Let \(\mc B\) be sequence-like, \(\mc A\subseteq\mc B\),
and \(\plI\notprewin G_{fin}(\mc A,\mc B)\). Then
\(\mc A\) is almost-sequence-like.
\end{proposition}
\begin{proof}
Let \(A\in\mc A\), and for all \(n<\omega\) let \(A_n=A\).
Then \(A_n\) is not a winning predetermined strategy for
\(\plI\), so \(\plII\) may choose finite subsets
\(B_n\subseteq A_n\) such that \(A_n'\bigcup\{B_n:n<\omega\}\in\mc B\subseteq\mc A\).

It follows that \(|A_n'|=\aleph_0\), and for any infinite subset
\(A_n''\subseteq A_n'\) (in particular, any cofinite subset),
\(A_n''\in\mc B\subseteq\mc A\). Thus \(\mc A\) is almost-sequence-like.
\end{proof}

\begin{corollary}
Let \(\mc B\) be sequence-like and \(\mc A\subseteq\mc B\).
Then \(\plI\win G_{fin}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{fin}(\mc A,\mc B)\).
\end{corollary}
\begin{proof}
Assuming \(\plI\notprewin G_{fin}(\mc A,\mc B)\), we have
\(\plI\notwin G_{fin}(\mc A,\mc B)\) by Proposition \ref{auto-asl}
and Theorem \ref{pawlikowskii}.
\end{proof}

>>>>

\begin{definition}
Say a collection \(\mc A\) is \term{\(\Gamma\)-like} if it satisfies the following
for each \(A\in\mc A\).
\begin{itemize}
\item \(|A|\geq\aleph_0\).
\item If \(A'\subseteq A\) and \(|A'|\geq\aleph_0\), then \(A'\in\mc A\).
\end{itemize}
\end{definition}

@@

It follows that \(\Gamma_X,\Gamma_{X,x}\) are both \(\Gamma\)-like. We also require
the following.

\begin{definition}
Say a collection \(\mc A\) is \term{almost-\(\Gamma\)-like} if
for each \(A\in\mc A\), there is \(A'\subseteq A\) such that:
\begin{itemize}
\item \(|A'|=\aleph_0\).
\item If \(A''\) is a cofinite subset of \(A'\), then \(A''\in\mc A\).
\end{itemize}
\end{definition}

So all \(\Gamma\)-like sets are almost-\(\Gamma\)-like.

\begin{theorem}
Let \(\mc B\) be \(\Gamma\)-like. Then \(\alpha_1(\mc A,\mc B)\) holds if and only
if \(\plI\notprewin G_{cf}(\mc A,\mc B)\).
\end{theorem}

@@

\begin{theorem}
Let \(\mc A\) be almost-\(\Gamma\)-like and \(\mc B\) be \(\Gamma\)-like. 
Then \(\alpha_2(\mc A,\mc B)\) holds if and only
if \(\plI\notprewin G_1(\mc A,\mc B)\).
\end{theorem}

@@

\begin{theorem}
Let \(\mc A\) be almost-\(\Gamma\)-like and \(\mc B\) be \(\Gamma\)-like. 
Then \(\alpha_4(\mc A,\mc B)\) holds if and only
if \(\plI\notprewin G_{<2}(\mc A,\mc B)\) if and only if
\(\plI\notprewin G_{fin}(\mc A,\mc B)\).
\end{theorem}

@@

\begin{theorem}
Let \(\mc B\) be \(\Gamma\)-like. Then \(\plI\prewin G_{<2}(\mc A,\mc B)\)
if and only if \(\plI\prewin G_{fin}(\mc A,\mc B)\).
\end{theorem}

\begin{proof}
Assume \(\bigcup\mc A\) is well-ordered.
Given a winning predetermined strategy \(A_n\) for \(\plI\) in
\(G_{<2}(\mc A,\mc B)\), consider \(F_n\in[A_n]^{<\aleph_0}\). We set
\[
  F_n^*
=
  \begin{cases}
    \emptyset
      & \text{ if }
    F_n\setminus\bigcup\{F_m:m<n\}=\emptyset
      \\
    \{\min(
      F_n\setminus\bigcup\{F_m:m<n\}
    )\}
      & \text{ otherwise}
  \end{cases}
\]
Since \(|F_n^*|<2\), we have that \(\bigcup\{F_n^*:n<\omega\}\not\in\mc B\).
In the case that \(\bigcup\{F_n^*:n<\omega\}\) is finite, we immediately
see that \(\bigcup\{F_n:n<\omega\}\) is also finite and therefore not
in \(\mc B\). Otherwise \(\bigcup\{F_n^*:n<\omega\}\not\in\mc B\)
is an infinite subset of \(\bigcup\{F_n:n<\omega\}\), and thus
\(\bigcup\{F_n:n<\omega\}\not\in\mc B\) too. Therefore
\(A_n\) is a winning predetermined strategy for \(\plI\) in
\(G_{fin}(\mc A,\mc B)\) as well.
\end{proof}

\begin{theorem}
Let \(\mc B\) be \(\Gamma\)-like. Then \(\plI\win G_{<2}(\mc A,\mc B)\)
if and only if \(\plI\win G_{fin}(\mc A,\mc B)\).
\end{theorem}

@@

We further note that the above proof technique could be
used to establish that perfect-information and
Markov winning strategies for \(\plII\) in
\(G_{fin}(\mc A,\mc B)\) may be improved to
be valid in \(G_{<2}(\mc A,\mc B)\), provided
\(\mc B\) is \(\Gamma\)-like. As such,
\(G_{<2}(\mc A,\mc B)\) and \(G_{fin}(\mc A,\mc B)\)
are effectively equivalent games under this hypothesis.
%\begin{theorem}
%Let \(\mc B\) be \(\Gamma\)-like. Then \(\plII\win G_{<2}(\mc A,\mc B)\)
%if and only if \(\plII\win G_{fin}(\mc A,\mc B)\).
%\end{theorem}

%\begin{theorem}
%Let \(\mc B\) be \(\Gamma\)-like. Then \(\plII\markwin G_{<2}(\mc A,\mc B)\)
%if and only if \(\plII\markwin G_{fin}(\mc A,\mc B)\).
%\end{theorem}

\begin{theorem}\label{pawlikowskii}
Let \(\mc A\) be almost-\(\Gamma\)-like and \(\mc B\) be \(\Gamma\)-like.
Then \(\plI\win G_{fin}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{fin}(\mc A,\mc B)\), and
\(\plI\win G_1(\mc A,\mc B)\) if and only if
\(\plI\prewin G_1(\mc A,\mc B)\).
\end{theorem}

\begin{proof}
We assume \(\plI\win G_{fin}(\mc A,\mc B)\)
and let the symbol \(\dagger\) mean \(<\aleph_0\)
(respectively, \(\plI\win G_1(\mc A,\mc B)\)
and \(\dagger=1\),
and for convenience we assume \(\plII\) plays
singleton subsets of \(\mc A\) rather than elements).
As \(\mc A\) is almost-\(\Gamma\)-like, there is a 
winning strategy \(\sigma\) where
\(|\sigma(s)|=\aleph_0\) and \(\sigma(s)\cap\bigcup\ran{s}=\emptyset\)
(that is, \(\sigma\) never replays the choices of \(\plII\))
for all partial plays \(s\) by \(\plII\).

For each \(s\in\omega^{<\omega}\), suppose 
\(F_{s\rest m}\in[\bigcup A]^{\dagger}\) 
is defined for each \(0<m\leq|s|\).
Then let \(s^\star:|s|\to[\bigcup\mc A]^{\dagger}\)
be defined by
\(s^\star(m)=F_{s\rest m+1}\), and define \(\tau':\omega^{<\omega}\to\mc A\)
by \(\tau'(s)=\sigma(s^\star)\). Finally, set 
\([\sigma(s^\star)]^{\dagger}=\{F_{s\concat\tuple{n}}:n<\omega\}\), and
for some bijection \(b:\omega^{<\omega}\to\omega\) let \(\tau(n)=\tau'(b(n))\)
be a predetermined strategy for \(\plI\) in \(G_{fin}(\mc A,\mc B)\)
(resp. \(G_1(\mc A,\mc B)\)).

Suppose \(\alpha\) is a counterattack by \(\plII\) against \(\tau\), so 
\[
  \alpha(n)
    \in
  [\tau(n)]^{\dagger}
    =
  [\tau'(b(n))]^{\dagger}
    =
  [\sigma(b(n)^\star)]^{\dagger}
\]
It follows that \(\alpha(n)=F_{b(n)\concat\tuple{m}}\) for some \(m<\omega\).
In particular, there is some infinite subset \(W\subseteq\omega\) and \(f\in\omega^\omega\)
such that \(\{\alpha(n):n\in W\}=\{F_{f\rest n+1}:n<\omega\}\).
Note here that \((f\rest n+1)^\star=(f\rest n)^\star\concat\tuple{F_{f\rest n+1}}\).
This shows that \(F_{f\rest n+1}\in[\sigma((f\rest n)^\star)]^{\dagger}\) 
is an attempt by \(\plII\) to defeat \(\sigma\), which fails. Thus 
\(\bigcup\{F_{f\rest n+1}:n<\omega\}=\bigcup\{\alpha(n):n\in W\}\not\in\mc B\),
and since this set is infinite (as \(\sigma\) prevents \(\plII\)
from repeating choices) we have \(\bigcup\{\alpha(n):n<\omega\}\not\in\mc B\) too.
Therefore \(\tau\) is winning.
\end{proof}

Note that the assumption in Theorem \ref{pawlikowskii} that \(\mc A\)
be almost-\(\Gamma\)-like cannot be omitted. In
[todo cite Clontz k-tactics in Gruenhage game]
an example of a space and point where \(\plI\win G_1(\mc A,\mc B)\) but
\(\plI\notprewin G_1(\mc A,\mc B)\) is given, where \(\mc A\) is
the set of open neighborhoods of the given point 
(which are all uncountable), 
and \(\mc B\) is the set of converging sequences to that point.
(Note that \(G_1(\mc A,\mc B)\) is called \(Gru_{O,P}(X,x)\) in that
paper. In fact, more is shown: \(\plI\) has a winning perfect-information
strategy, but any strategy that only uses 
the most recent \(k\) moves of \(\plII\) and the round number
can be defeated, where \(k\) is any natural number.)

\begin{proposition}\label{auto-asl}
Let \(\mc B\) be \(\Gamma\)-like, \(\mc B\subseteq\mc A\),
and \(\plI\notprewin G_{fin}(\mc A,\mc B)\). Then
\(\mc A\) is almost-\(\Gamma\)-like.
\end{proposition}
\begin{proof}
Let \(A\in\mc A\), and for all \(n<\omega\) let \(A_n=A\).
Then \(A_n\) is not a winning predetermined strategy for
\(\plI\), so \(\plII\) may choose finite sets
\(B_n\subseteq A_n=A\) such that 
\(A'=\bigcup\{B_n:n<\omega\}\in\mc B\subseteq\mc A\).

It follows that \(A'\subseteq A\) and \(|A'|=\aleph_0\), 
and for any infinite subset
\(A''\subseteq A'\) (in particular, any cofinite subset),
\(A''\in\mc B\subseteq\mc A\). Thus \(\mc A\) is almost-\(\Gamma\)-like.
\end{proof}

\begin{corollary}
Let \(\mc B\) be \(\Gamma\)-like and \(\mc B\subseteq\mc A\).
Then \(\plI\win G_{fin}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{fin}(\mc A,\mc B)\), and
\(\plI\win G_{1}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{1}(\mc A,\mc B)\).
\end{corollary}
\begin{proof}
Assuming \(\plI\notprewin G_{fin}(\mc A,\mc B)\), we have
\(\plI\notwin G_{fin}(\mc A,\mc B)\) by Proposition \ref{auto-asl}
and Theorem \ref{pawlikowskii}.

Similarly, assuming \(\plI\notprewin G_{1}(\mc A,\mc B)
\Rightarrow\plI\notprewin G_{fin}(\mc A,\mc B)\), we have
\(\plI\notwin G_{1}(\mc A,\mc B)\) by Proposition \ref{auto-asl}
and Theorem \ref{pawlikowskii}.
\end{proof}

This corollary generalizes e.g. Theorems 26 and 30 of [cite Scheepers 1996 Ramsey]
and Theorem 5 of [cite MR2119791].
